More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-26 2016-Weak-FH-Bisim/WeakBisim.tex
correction + ajout comments

For example $a^{i\in\{3\}}$ is
the mapping with a single entry $a$ at index $3$; exceptionally, such mappings with
only a few entries will also be denoted $(3\mapsto a)$.
When this is not ambiguous, we shall use abusive vocabulary and
notations for sets, and typically write ``indexed set over I'' when  
formally we should speak of multisets, and ``$x\in
A_i^{i\in I}$'' to mean $\exists i\in I.\, x=A_i$.
An empty family is denoted $[]$. % (it can be defined as $a_i^{i\in\emptyset}$).
To simplify equations, an indexed set can be denoted $\set{M}$
instead of $M_l^{l\in L}$ when $L$ is irrelevant.

@@

 Then 
$\mathcal{R}$ is an FH-bisimulation iff for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
	\TODO{change fig: s1,s2 $\to$ s,t}
\end{minipage}

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNets $\pNet'$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNet'\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNet'\},\Sort_j^{j\in J\setminus \{j_0\}},
	\overline{\symb{SV}}\myrangle
	\]
\end{definition}

@@

       	$\Pred_{s,p} \land \Pred_{OT}\\
       	\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       	\left( \forall j. b_j=b_{jpx}  \Rightarrow \Pred_{OT_{px}}
       	\land v\!=\!v_{px} \land
       	\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)$

       	for any open transition $OT_{px}$, since
       	$(t_p,u|\Pred_{p,u})\in\mathcal{R}'$ there exist open transitions
       	$OT_{pxy}^{y\in Y} \subseteq \mathcal{T}_3$:

@@

       	\end{mathpar}
       	such that  $\forall y, J_{px}=J_{pxy}, 
       	(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$; and  \\
       	$\Pred_{p,u} \land \Pred_{OT_{p x}}\\
       	\hspace{1cm} \implies\!\!\! \bigvee_{y\in Y}
       	\left( \forall j. b_{jpx}=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
       	\land v_{px}\!=\!v_{pxy} \land
       	\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\right)$.

@@

       	First, $\forall y, \forall x, \forall p, J_1=J_{px}=J_{pxy}, 
       	({s}',u_{pxy}|\Pred\,'_{{pxy}})\in \mathcal{R}''$. 
       	Indeed for any 
       	$p$, 
       	$x$, and 
       	$y$, $t_{px}$
       	relate ${s}'$ and $u_{pxy}$, indeed
       	$(s',t_{px}|\Pred_{{px}})\in \mathcal{R}$
       	and $(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$. 
       	More precisely, $t_{px} \in ({t'_p})^{p\in P'}$ where $({t'_p})^{p\in 
       		P'}$ \TODO{Pas sure prime utile t, or p, or P ; SI:  $P\neq P'$ et 
       		les $t_p'$ sont pas les $t_p$}
       	is 
       	the set of states relating ${s}'$ and $u_{pxy}$.
       	Consequently, for all $p$, $x$, $y$, $\Pred_{px}\land 
       	\Pred_{{pxy}}\implies 
       	\Pred\,'_{{pxy}}$ (this is one element of the  disjunction defining the 
       	predicate 
       	relating ${s}'$ and $u_{pxy}$ in the definition of $\mathcal{R}''$).

@@

       	Second, by definition of bisimulation we need (recall that $\Pred_{s,u}$ is the 
       	original predicate relating $s$ and $u$ by definition of the transitive 
       	closure):\\
       	\begin{small}
       		$\Pred_{s,u} \land \Pred_{OT}\\
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
       		\left( \forall j. b_j=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
       		\land v\!=\!v_{pxy} \land
       		\Pred\,'_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}}\right)$.
       	\end{small}

       	We have, for all $p$
       	\begin{small}
       		$\Pred_{s,p} \land \Pred_{OT}\land\Pred_{p,u}\\
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\left( \forall j. b_j=b_{jpx}  \Rightarrow \Pred_{OT_{px}}
       		\land v\!=\!v_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)\land\Pred_{p,u}\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\left( \forall j. b_j=b_{jpx}  \Rightarrow (\Pred_{OT_{px}}\land\Pred_{p,u})
       		\land v\!=\!v_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\big( \forall j. b_j=b_{jpx}  \Rightarrow (\bigvee_{y\in Y} 
       		\big( \forall j'. b_{j'px}=b_{j'pxy}  \Rightarrow \Pred_{OT_{pxy}}
       		\land v_{px}\!=\!v_{pxy}\\\hspace{3em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\big))
       		\land v\!=\!v_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\big)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X} \bigvee_{y\in Y}
       		\big( \forall j, j'. b_j=b_{jpx} \land b_{j'px}=b_{j'pxy}
       		\Rightarrow \big( 
       		\Pred_{OT_{pxy}}
       		\land v\!=v_{px}\!=\!v_{pxy}\\\hspace{3em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}
       		\land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\big)\big)
       		$

@@

By construction, the Post substitutions only have an effect on the  
variables of the open automaton they belong to, they also produce terms containing only 
variables of the open automaton they belong to. Finally, because of the domain of the 
substitutions  and of the predicates, we have:\\
       	$\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\land 
       	\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
       	= 
       	\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\land
       	\Pred_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
       	\implies\Pred\,'_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
       	= \Pred\!'_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}} $

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I},
				\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {v}
				\ostate{s_i^{\prime i\in L}}}
		}
	\]
		with $Q$ ``moving'' (i.e. $J\cap\Holes(Q)\neq\emptyset$ or $I\cap L_Q\neq 
		\emptyset$).
		 Then there exist $v_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\setminus \Holes(Q) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {v}
				\ostate{s_i'^{\, i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
				\ostate{s_i'^{\, i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred\subst{v_Q\leftarrow b_{j_0}} =(\Pred\,'
		\land \Pred\,'')$, $\Post=\Post'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

   \begin{small}\begin{proof}
       Consider each premise of the open transition (as constructed by rule TR2 in Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P[Q]_{j_0}$ and try to prove the equivalent premise for 
$P$. 
First, $K$ and the synchronisation vector $SV_k$ are unchanged\footnotemark (however 
$j_0$ passes from 
the set of subnets to the set of holes). Then $SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. 
Now focus on OTs of the subnets\footnotemark[\thefootnote]:\\[-2ex]
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
	\ostate{s_{i}^{\prime\ i \in L_m}}}\]
Only elements of $I_k$ are useful to assert the premise for reduction of $P$; the last 
one ensures (note that $Q$ is at place $j_0$, and by definition of the open transition 
for $P[Q]_{j_0}$, $I_{j_0}=I\cap L_Q$, 
$L_{j_0}=L_Q$):\\[-2ex]
	\[Q\models{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {v_{j_0}}
			\ostate{s_i'^{\, i\in L_Q}}}
	}\]
This already ensures the second part of the conclusion if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Now let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J_k 
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$; the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m 
\land \Pred(SV,a_i^{i\in I_k},b_j^{j\in J_k\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred(SV,a_i^{i\in I_k},b_j^{j\in 
J_k},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J_k\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'_k 
$. Modulo renaming of fresh variables, this is identical to the predicate that occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \\\\
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {v}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~and~}
%\[
Q\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I_Q},
		\{\xrightarrow{b_j}_j\}^{j\in J_Q},\\\\
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I \uplus I_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\setminus\{j_0\} \uplus J_Q}, 
			\Pred\subst{b_{j_0}\leftarrow v_Q}\land\Pred\,',  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {v}
			\ostate{s_i^{\prime i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\begin{small}

>>>>

For example $a^{i\in\{3\}}$ is
the mapping with a single entry $a$ at index $3$; exceptionally, such mappings with
only a few entries will also be denoted $(3\mapsto a)$. \TODO{We need to define update as $A \uplus \{j \mapsto a\}$ or $A[j \mapsto a]$}
When this is not ambiguous, we shall use abusive vocabulary and
notations for sets, and typically write ``indexed set over I'' when  
formally we should speak of multisets, and ``$x\in
A_i^{i\in I}$'' to mean $\exists i\in I.\, x=A_i$.
An empty family is denoted $[]$. % (it can be defined as $a_i^{i\in\emptyset}$).
To simplify equations, an indexed set can be denoted $\set{M}$
instead of $M_l^{l\in L}$ when $L$ is irrelevant.

@@

 Then 
$\mathcal{R}$ is an FH-bisimulation iff for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
	\TODO{change fig: s1,s2 $\to$ s,t. DONE we will check if it is OK. Why did we use tilde notation for R}
\end{minipage}

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNets $\pNet'$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNet'\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNet'\},\Sort_j^{j\in J\setminus \{j_0\}},
	\overline{\symb{SV}}\myrangle
	\]
\end{definition}
\TODO{$\uplus\{j_0\mapsto \pNet'\}$ non defini}

@@

       	$\Pred_{s,p} \land \Pred_{OT}\\
       	\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       	\left( \forall j. b_j=b_{jpx}  \Rightarrow \Pred_{OT_{px}}
       	\land v\!=\!v_{px} \land
       	\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)$\\

For any open transition $OT_{px}$, since
       	$(t_p,u|\Pred_{p,u})\in\mathcal{R}'$ there exist open transitions
       	$OT_{pxy}^{y\in Y} \subseteq \mathcal{T}_3$:

@@

       	\end{mathpar}
       	such that  $\forall y, J_{px}=J_{pxy}, 
       	(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$; and  \\
       	$\Pred\,'_{p,u} \land \Pred_{OT_{p x}}\\
       	\hspace{1cm} \implies\!\!\! \bigvee_{y\in Y}
       	\left( \forall j. b_{jpx}=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
       	\land v_{px}\!=\!v_{pxy} \land
       	\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\right)$\\.

@@

       	First,  $\forall y, \forall x, \forall p, J_1=J_{px}=J_{pxy}, 
       	({s}',u_{pxy}|\Pred\,'_{{pxy}})\in \mathcal{R}''$. 
       	Indeed for any 
       	$p$, 
       	$x$, and 
       	$y$, $t_{px}$
       	relate ${s}'$ and $u_{pxy}$, indeed
       	$(s',t_{px}|\Pred_{{px}})\in \mathcal{R}$
       	and $(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$. 
       	More precisely, $t_{px} \in ({t'_p})^{p\in P'}$ where $({t'_p})^{p\in 
       		P'}$ is 
       	the set of states relating ${s}'$ and $u_{pxy}$.
       	Consequently, for all $p$, $x$, $y$, $\Pred_{px}\land 
       	\Pred_{{pxy}}\implies 
       	\Pred\,'_{{pxy}}$ (this is one element of the  disjunction defining the 
       	predicate 
       	relating ${s}'$ and $u_{pxy}$ in the definition of $\mathcal{R}''$).

@@

       	Second, by definition of bisimulation we need (recall that $\Pred_{s,u}$ is the 
       	original predicate relating $s$ and $u$ by definition of the transitive 
       	closure):\\
       	\begin{small}
       		$\Pred_{s,u} \land \Pred_{OT}\\
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
       		\left( \forall j. b_j=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
       		\land v\!=\!v_{pxy} \land
       		\Pred\,'_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}}\right)$.
       	\end{small}
\TODO{From we will add notation (1) and (2)}        	
       	We have, for all $p$
       	\begin{small}
       		$\Pred_{s,p} \land \Pred_{OT}\land\Pred_{p,u}\\
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\left( \forall j. b_j=b_{jpx}  \Rightarrow \Pred_{OT_{px}}
       		\land v\!=\!v_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)\land\Pred_{p,u}\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\left( \forall j. b_j=b_{jpx}  \Rightarrow (\Pred_{OT_{px}}\land\Pred_{p,u})
       		\land v\!=\!v_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\big( \forall j. b_j=b_{jpx}  \Rightarrow (\bigvee_{y\in Y} 
       		\big( \forall j'. b_{j'px}=b_{j'pxy}  \Rightarrow \Pred_{OT_{pxy}}
       		\land v_{px}\!=\!v_{pxy}\\\hspace{3em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\big))
       		\land v\!=\!v_{px} \land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\big)\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X} \bigvee_{y\in Y}
       		\big( \forall j, j'. b_j=b_{jpx} \land b_{j'px}=b_{j'pxy}
       		\Rightarrow \big( 
       		\Pred_{OT_{pxy}}
       		\land v\!=v_{px}\!=\!v_{pxy}\\\hspace{3em}~ \land
       		\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}
       		\land
       		\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\big)\big)
       		$

@@

By construction, four substitutions $\subst{~}$ only have an effect on the  
variables of the open automaton they belong to, they also produce terms containing only 
variables of the open automaton they belong to. Finally, because of the domain of the 
substitutions  and of the predicates, we have:\\
       	$\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\land 
       	\Pred_{{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
       	\Leftrightarrow % we changed =
       	\Pred_{{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\land
       	\Pred_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
       	\implies\Pred\,'_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
       	\Leftrightarrow % we changed = 
       	 \Pred\!'_{{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}} $\\

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I},
				\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {v}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with $Q$ ``moving'' \TODO{or is reduced?}, i.e. $J\cap\Holes(Q)\neq\emptyset$ or $I\cap L_Q\neq 
		\emptyset$.
		 Then there exist $v_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {v}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{\,i\in I\cap L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred\subst{v_Q\leftarrow b_{j_0}} =(\Pred\,'
		\land \Pred\,'')$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

   \begin{small}\begin{proof}
       Consider each premise of the open transition (as constructed by rule TR2 in Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P[Q]_{j_0}$ and try to prove the equivalent premise for 
$P$. 
First, $K$ and the synchronisation vector $SV_k$ are unchanged\footnotemark (however 
$j_0$ passes from 
the set of subnets to the set of holes). Then $SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. 
\TODO{Revoir la def des leaves set ou multisets}

Now focus on OTs of the subnets\footnotemark[\thefootnote]:\\[-2ex]
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
	\ostate{s_{i}'^{\, i \in L_m}}}\]
Only elements of $I_k$ are useful to assert the premise for reduction of $P$; the last 
one ensures (note that $Q$ is at place $j_0$, and by definition of the open transition 
for $P[Q]_{j_0}$, $I_{j_0}=I\cap L_Q$, 
$L_{j_0}=L_Q$):\\[-2ex]
	\[Q\models{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {v_{j_0}}
			\ostate{s_i'^{\, i\in L_Q}}}
	}\]
\TODO{We need to add: Note that $J_{j_0}=	J\cap\Holes(Q)$ this is a direct consequence of Tr2??}

This already ensures the second part of the conclusion if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Now let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J_k 
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$; the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m 
\land \Pred(SV,a_i^{i\in I_k},b_j^{j\in J_k\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred(SV,a_i^{i\in I_k},b_j^{j\in 
J_k},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J_k\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'_k 
$. Modulo renaming of fresh variables, this is identical to the predicate that occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \\\\
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {v}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~and~}
%\[
Q\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I_Q},
		\{\xrightarrow{b_j}_j\}^{j\in J_Q},\\\\
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I \uplus I_Q},
			\{\xrightarrow{b_j}_j\}^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\subst{b_{j_0}\leftarrow v_Q}\land\Pred\,',  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {v}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\begin{small}
